---
title: "Getting Started with Benvos"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with Benvos}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{ggplot2}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette serves to introduce the philosophy and structure of the `rbenvo` package, which impliments `Benvo` classes: Built Environment Objects. These classes make working with built environment data easier by offering a common interface to do so. We'll illustrate some of the very basics of how to use benvo's in this vignette as well
as give some references as to how it might be used in other packages or settings.

We'll begin by loading the library and the two example datasets involving subjects living near Fast Food Restaurants (FFR). One of the two datasets will be for the subject level information, `FFR_subjects`, the other will 
contain some measure of exposure (space/time) that describe how long and how close subjects are located near the nearby Built Environment Features (BEFs).

```{r setup,message=F,results=F}
library(rbenvo)
library(ggplot2)
theme_set(theme_bw())
data("FFR_distances")
data("FFR_subjects")
```


The subject dataframe should look familiar. This is a tidy dataset with an ID and two subject-level measures: BMI and sex.

```{r subjdf}
head(FFR_subjects,5)
```

The distance dataframe is a tidy "long" dataframe, containing multiple nearby BEF exposures per subject. Note that there *must be* at least two columns in every BEF dataset added to a benvo. One for the id and one for the Distance/Time data measured. Additionally, the latter columns must be named as "Distance" or "Time",as benvo's use these titles to perform their methods.
```{r befdf}
head(FFR_distances,5)
```

In order to create a benvo we'll pass the two dataframes to the `benvo()` function, wrapping the FFR_distances data frame in a list. 

```{r mkbenvo}
bdf <- benvo(subject_data = FFR_subjects,
             bef_data = list(FFR=FFR_distances))
bdf
```
## Simple Descriptives

We can see from the above a simple summary of the kind of feature that we have data on. If we call the `bef_summary` function we'll get a more detailed understanding of the bef data available.

```{r summary}
bef_summary(bdf)
```

Finally, as one last descriptive offering, we can plot a point-range of the distances for each subject using the 
`plot` function and passing the BEF label we want plotted. You'll see that there are several observations missing, since not every subject lives near a FFR.


```{r plot}
plot(bdf)
```


# Other packages

While the above descriptive statistics are nice, the real power of using `benvos` is that they offer 
functions like `joinvo` to easily join the subject and bef data frames, which can be useful in setting up the data for models in other packages. Below we'll use the `joinvo` function, passing the benvo, bef character label, and `TRUE` to the tibble argument, denoting that we want a tibble object returned from the function.

```{r}
joinvo(bdf,"FFR",tibble=T)
```

Similarly, it is often the case - in my work at least - that some sort of model involving both the built environment data and the subject level data is needed. This means that information is incorporated at the subject level. This could manifest in, for example, the construction of 
a subject design matrix. Below I'll demonstrate how this works in a cross sectional as well as longitudinal framework, whilst highlighting how some of the previous functions take advantage
of the longitudinal structure.

```{r subjectdesign_example}
str(subject_design(bdf,BMI ~ sex))
```

### Longitudinal Example

For our longitudinal example we'll look at a dataset that imagines a hypothetical set of grocery or "Healthy Food Stores"(HFS)  near subjects. Subjects or grocery stores may move over
time so this will be reflected in our data. Below we'll repeat the similar descriptives from above, highlighting the differences in a longitudinal `benvo`.
```{r longitudinal_example}
data("longitudinal_HFS")
longitudinal_HFS
```

```{r longitudinal_summary}
bef_summary(longitudinal_HFS)
```


```{r long_plot}
plot(longitudinal_HFS,term ="HFS",component = "Distance", p = 0.9)
```

Analogous to `subject_design` there is also the `longitudinal_design` function which effectively allows for `lme4` style formulas to be used, incorporating subject or group level effects from the `subject_data` in the `benvo` and returning a `glmod` object along with the corresponding `X` and `y` design matrix and outcome vector, respectively.

```{r longitudinal_design_ex}
str(longitudinal_design(longitudinal_HFS,BMI ~ sex + (1|id)),max.level=1)
```

## Further Information

This completes this preliminary introduction to the `rbenvo` package. For current uses please see the [`rsstap`](https://apeterson91.github.io/rsstap/) or  [`rstapDP`](https://apeterson91.github.io/rstapDP/) packages. I've found that having these data structures greatly simplifies my code and allows for data to be organized more easily. If you're working with similar data, I hope you find the same!


I'm hoping to add more functions and classes soon that will make it easy to construct benvos from latitude-longitude coordinates or `sf` BEF data structures. Stay tuned for these updates.
